A comparison on R tools for nonlinear least squares modeling

### 2023-03-14 RefactoringNLS/ComparingNLStools.Rmd
cleanup

# Abstract

@@

There are a number of other CRAN packages worth commentary, and we will give
some capsule comments later. Packages outside of CRAN, even if they might be
worthy of review, will not be included here.

??move down
## Other CRAN packages

`nlstools` (@p-nlstools) -- Several tools for assessing the quality of fit of a gaussian nonlinear model are provided.

nlsmsn -- Fit univariate non-linear scale mixture of skew-normal(NL-SMSN) regression, details in Garay, Lachos and Abanto-Valle (2011)

`nlstac` (@m-nlstac) -- Set of functions implementing the algorithm described in Fernandez 
    Torvisco et al. (2018) for fitting separable nonlinear regression curves.

    ?? check ref and add to bib.??

easynls -- Fit and plot some nonlinear models. ?? no vignette!

nlraa -- a set of nonlinear *selfStart* models. Most include analytic Jacobian
code. (@MiguezNLRAA2021)

optimx -- not really a nonlinear least squares method, but has general optimizers
that can be applied to minimize a nonlinear function e.g., a nonlinear sum of 
squares.

It is likely that there are a number of other tools that could be relevant to
this comparison in the Bioconductor (@Bioconductor) collection 
and on repositories like
Github (https://github.com) and Gitlab (https://about.gitlab.com), but we will 
not pursue those possibilities here.

@@

?? weighting, indexed parameters, partially linear ...

@@

?? presence of vignettes, demos, (good?) examples

>>>>

## Abstract

@@

There are a number of other CRAN packages worth commentary, and we will give
some capsule comments later. It is likely that there are a number of other 
tools that could be relevant to
this comparison in the Bioconductor (@Bioconductor) collection 
and on repositories like
Github (https://github.com) and Gitlab (https://about.gitlab.com), but we will 
not pursue those possibilities here.

@@

Users often have particular problems that introduce other requirements. It is
common to weight observations where their reliability is thought to differ.
Some situations present a number of related situations where a combined model
with some common and some local parameters are present, and the local parameters
could be indexed. The common exponential decay and related **partially linear**
models (sometimes called **separable** models) is a difficult one that can 
benefit from special treatment.

@@

We believe that the availability of good documentation with examples that users
can modify to deal with their own problems is essential. The presence of vignettes, 
demos, and tests indicates a package is well-supported.

## Other CRAN packages

`nlstools` (@p-nlstools) -- This package provides several tools for aiding the estimation of nonlinear models, particularly
using `nls()`. The vignette is actually a journal article, and the authors have considerable experience in the subject.

`nlsmsn` (@m-nlsmsn) -- Fit univariate non-linear scale mixture of skew-normal(NL-SMSN) regression, with details in Garay, Lachos and Abanto-Valle (2011). The problem here is to minimize an objective which is modified from the traditional sum of squared residuals.

`nlstac` (@m-nlstac) -- Set of functions implementing the algorithm described in Fernandez 
    Torvisco et al. (2018) for fitting separable nonlinear regression curves. The special 
    class of problem for which this package is intended is an important and difficult one.
    No vignette is provided, unfortunately.

`easynls`  -- Fit and plot some nonlinear models. Thirteen models are treated, but there is minimal 
documentation and no vignette. Package `nlraa` is to be preferred.

`nlraa`  (@MiguezNLRAA2021) -- a set of nonlinear *selfStart* models. Most include analytic Jacobian
code.

`optimx` (@p-optimx) -- not really a nonlinear least squares method, but has general optimizers
that can be applied to minimize a nonlinear function e.g., a nonlinear sum of 
squares, as well as other objectives. This is not generally recommended if nonlinear least 
squares programs can be easily used, but provides a way to approach problems when others are
unavailable.
